---
html_document: default
author: "Sanjay Pillay & Angelo Bravo"
date: "10/9/2019"
output: 
  html_document: 
    fig_width: 8
title: "Case Study 1"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
The case study involves a exploratory data analysis of breweries and the beer they manufacture in all the states of US. The study finally ends with some recommendations for the study sponsores (CEO and CFO of Budweiser) as to where to target their next product and the reasoning behind those recommendations.

# Video Presentation
https://youtu.be/gfOp74IlmpQ

# Shinny App
https://dds-app1.shinyapps.io/shinnyunit12/

Importing Libraries
```{r message=FALSE}
#uncomment below installpks command if running for the first time
#install.packages("dplyr", "naniar","mice","VIM","stringi","stringr"."rvest","purrr","tidyverse","tidyr","ggthemes","plotly","ggplot2","reshape2","GGally","caret","class")
library(dplyr) #join etc
library(naniar) # check nulls
library(mice) # imputing
library(VIM) # view imputed datas
library(stringi)
library(stringr)
library(rvest) #html_table, html_node
library(purrr)
library(tidyverse) # Data cleaning
library(tidyr) # Data cleaning
library(ggthemes) #Plotting
library(plotly) #Plotting
library(ggplot2) #Plotting
library(reshape2) # melt
library(GGally) # ggpairs
library(caret) #Confution matrix
library(class)
library(caret)
library(e1071)
```

Importing data and filling in missing data with mice package
```{r}
#Read supplied data
cdw = getwd()
brewwriesData = read.csv(paste(getwd(),"/data/","Breweries.csv", sep = ""),header = TRUE)
beerData = read.csv(paste(getwd(),"/data/","Beers.csv", sep = ""),header = TRUE)
#Analyse the data
gg_miss_var(beerData)
# about 58.3% of data are not missing any values
# 40% of IBU and 2.6% of ABV are misisng values

#Plot the missing value using mice package
aggr_plot <- aggr(beerData, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(beerData), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

# left box plot to distribution of AVB with and without misisng IBU ,Right shows there are no IBU observations where ABV is missing
marginplot(beerData[c('IBU','ABV')])

#Summary data
summary(brewwriesData)
summary(beerData)

# Deal with missing values
# We will get all missing vars to build the predictorMatrix to be passed to mice
missVars <- names(beerData)[colSums(is.na(beerData)) > 0]

#Get all the variables names in the dataset
allVars <- names(beerData)
#Code borrwoed from https://rpubs.com/kaz_yos/mice-exclude
#Initialize the matrix witl all row  column having th evar names form above
predictorMatrix <- matrix(0, ncol = length(allVars), nrow = length(allVars))
rownames(predictorMatrix) <- allVars
colnames(predictorMatrix) <- allVars

#List the variables we want to be used for the calculations
imputerVars <- c("ABV","IBU","Style")

## Keep variables that actually exist in dataset
imputerVars <- intersect(unique(imputerVars), allVars)
imputerMatrix <- predictorMatrix
imputerMatrix[,imputerVars] <- 1

#Specify variables to be imputed
imputedOnlyVars <- c("ABV","IBU")
imputedVars <- intersect(unique(c(imputedOnlyVars, imputerVars)), missVars)

imputedMatrix <- predictorMatrix
imputedMatrix[imputedVars,] <- 1

predictorMatrix <- imputerMatrix * imputedMatrix

## Diagonals must be zeros (a variable cannot impute itself)
diag(predictorMatrix) <- 0

#Matrix to feed mice call
predictorMatrix

#Generate 5 sets using 50 iterations using pnm (Predictive mean matching) method
#imputedBeer = mice(beerData,m=5,maxit=50,meth='pmm',seed=500, predictorMatrix = predictorMatrix)
```

```{r echo=FALSE, warning=FALSE,message=FALSE, results='hide'}

imputedBeer = mice(beerData,m=5,maxit=50,meth='pmm',seed=500, predictorMatrix = predictorMatrix)

```


Visualizng Imputed Dataset
```{r}
#mice density polt
densityplot(imputedBeer,IBU~ABV|.imp)
completedData = mice::complete(imputedBeer,1)
xyplot(imputedBeer,ABV ~ IBU)
xyplot(beerData$ABV ~ beerData$IBU,data =beerData)

###1 How many Brueries in each state
brewwriesData %>%  count(State)
beerBrewries = merge(completedData, brewwriesData, by.x = "Brewery_id", by.y = "Brew_ID")
head(beerBrewries,6)
tail(beerBrewries,6)

## Media Alcohol/Bitterness for each state
abByState = beerBrewries %>% group_by(State) %>% summarize(medianIBU = median(IBU), medianABV = median(ABV))
abByState = abByState %>% mutate(Bitterness = 'Middle') 
abByState[which.max(abByState$medianIBU),]$Bitterness = "Max"
abByState[which.min(abByState$medianIBU),]$Bitterness = "Min"
abByState = abByState %>% mutate(Alcohol = 'Middle') 
abByState[which.max(abByState$medianABV),]$Alcohol = "Max"
abByState[which.min(abByState$medianABV),]$Alcohol = "Min"

#Comparision bar chart per state
p =ggplot(abByState,aes(x = State,y = medianIBU)) + 
    geom_bar(aes(fill = Bitterness),stat = "identity") +
  labs(title="Median bitterness by state", y = "IBU", x="State" ) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p + theme(legend.position = "top")
p =ggplot(abByState,aes(x = State,y = medianABV*100)) + 
    geom_bar(aes(fill = Alcohol),stat = "identity") +
  labs(title="Median alcohol % content by state", y = "ABV %", x="State" ) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p + theme(legend.position = "top")

#Identify beer Categories by IPA//Ale/Other
beerBrewries = beerBrewries %>% mutate(bClass = 'Other') 
beerBrewries = beerBrewries %>% mutate(ABVpercent = ABV*100) 
beerBrewries[grepl('IPA',beerBrewries$Style, ignore.case = TRUE),]$bClass = "IPA"
beerBrewries[grepl('Ale',beerBrewries$Style, ignore.case = TRUE),]$bClass = "Ale"
beerBrewries = beerBrewries %>% mutate(bClass = as.factor(bClass)) 

###6 ABV statistics
summary(beerBrewries$ABVpercent)
beerBrewries %>% ggplot(aes(x = bClass, y = ABVpercent)) + geom_boxplot() + labs(title="ABV % distribution by beer category", y = "ABV %", x="Beer Category" ) 
beerBrewries %>% 
ggplot(aes(x = log(ABVpercent), color = bClass)) + geom_histogram() + facet_grid(rows = vars(bClass))+ labs(title="ABV % distribution by beer category", y = "Count", x="% ABV" ) +theme_classic()

###7 relation between ABV and IBU
beerBrewries %>% ggplot(aes(x = IBU, y = ABV)) + geom_point(aes(colour = bClass)) + labs(title="Relation of ABV vs IBU after including missing data", y = "ABV", x="IBU" ) 

beerBrewries %>%  select( IBU, ABVpercent,bClass) %>% ggpairs(aes(color = bClass))
```

Preparing dataset for KNN
Creating IPA/Ale only dataframe
```{r}
#### Q 8 & 9
#Set up data for Training and test set 50/50
ipa_ale_df <- beerBrewries %>% filter(bClass %in% c("IPA", "Ale"))
ipa_df <- beerBrewries %>% filter(bClass %in% c("IPA"))
ale_df <- beerBrewries %>% filter(bClass %in% c("Ale"))
ipa_ale_df$bClass <-  droplevels(ipa_ale_df$bClass)
ipa_ale_df$bClass <- factor(ipa_ale_df$bClass, levels = c("IPA", "Ale"))
levels(ipa_ale_df$bClass)
#removing ale,other levels
ipa_df$bClass <- droplevels(ipa_df$bClass)
#adding ale level to ipa_df
ipa_df$bClass <- factor(ipa_df$bClass, levels = c(levels(ipa_df$bClass), "Ale"))
#same process to ale_df
ale_df$bClass <- droplevels(ale_df$bClass)
#adding ale level to ipa_df
ale_df$bClass <- factor(ale_df$bClass, levels = c(levels(ale_df$bClass), "IPA"))
```

Fitting 100 models for 100 Ks and getting the mean accuracy for each K
We are trying to show that ABV and IBU is significantly different for IPAs and Ales
by showing you can get more than 50% accuracy on a 50/50 split training set of IPAs and
Ales. 
```{r}
#IPA,ALE VS IPA,ALE (50/50 split)
set.seed(100)
n = 1
accuracydf <- data.frame(accuracy = numeric(10000), k = numeric(10000), 
                         sensitivity = numeric(10000), specificity = numeric(10000))
for(i in 1:100) {
  #Get dataset with evenly distributed IPAs and Ales (50/50)
  ipa_test_ind <- sample(1:nrow(ipa_df), nrow(ipa_df)-250)
  ale_df_ind <- sample(1:nrow(ale_df), 250)
  #building train set
  ipa_ale_train <- rbind(ipa_df[-ipa_test_ind,], ale_df[ale_df_ind,])
  ipa_ale_overall_test <- ipa_ale_df[sapply(ipa_ale_df$Name.x, function(x) x %in% ipa_ale_train$Name.x) == FALSE,]
  ipa_ale_test_ind <- sample(1:nrow(ipa_ale_overall_test), round(.2 * nrow(ipa_ale_train)))
  ipa_ale_test <- ipa_ale_overall_test[ipa_ale_test_ind,]
  
  
  #storing accuracy data in data frame
  for(j in 1:100) {
    classifications <- knn(ipa_ale_train[c("ABV", "IBU")], ipa_ale_test[c("ABV", "IBU")], cl = ipa_ale_train$bClass, k = j, prob = F)
    CM = confusionMatrix(table(classifications, ipa_ale_test$bClass))
    accuracydf$accuracy[n] = CM$overall[1]
    accuracydf$sensitivity[n] = CM$byClass["Sensitivity"]
    accuracydf$specificity[n] = CM$byClass["Specificity"]
    accuracydf$k[n] = j
    n = n + 1
  }
}
max(accuracydf$accuracy, na.rm = TRUE)

#accuracydf[accuracydf$accuracy == 0.91,]
summary_acc_df <- accuracydf %>% group_by(k) %>% summarise(mean_accuracy = mean(accuracy), 
                                                           mean_sensitivity = mean(sensitivity),
                                                           mean_specificity = mean(specificity, na.rm = T))
overall_mean_accuracy <- summary_acc_df[which.max(summary_acc_df$mean_accuracy),]
#overall_mean_accuracy
acc_df <- as.data.frame(summary_acc_df)
accuracydf %>% ggplot(aes(x=k, y = accuracy)) + geom_point() + geom_smooth()
str(acc_df)
library(ggthemes)
acc_df %>% ggplot(aes(x = k, y = mean_accuracy)) + geom_point() + geom_smooth() + theme_economist() + 
  ggtitle("Mean Accuracy For 100 Iterations vs. K") + xlab("K") + ylab("Mean Accuracy")
```


We will now get ABV confidence intervals for IPAs and Ales
```{r}
#run t-test two test if our overall_mean_accuracy is greater than 50
#t.test
#t.test(ipa_ale_df$ABV[ipa_ale_df$bClass=="IPA",], ipa_ale_df)
plot(density(ipa_ale_df$ABV[ipa_ale_df$bClass=="IPA"]))
hist(ipa_ale_df$ABV[ipa_ale_df$bClass=="IPA"], col=rgb(1,0,0,0.5), xlim = range(.01,.15), main="Ale and IPA ABV Histogram",
     xlab = "ABV", ylab = "Count")
hist(ipa_ale_df$ABV[ipa_ale_df$bClass=="Ale"], col=rgb(0,0,1,0.5), add=T)
#run a two-sample to test if the groups have significantly different ABVs
t.test(ipa_ale_df$ABV[ipa_ale_df$bClass=="IPA"], ipa_ale_df$ABV[ipa_ale_df$bClass=="Ale"])
t.test(ipa_ale_df$IBU[ipa_ale_df$bClass=="IPA"], ipa_ale_df$IBU[ipa_ale_df$bClass=="Ale"])
#run a ttest on each group in order to determine a 95% CI of abv for each group
t.test(ipa_ale_df$ABV[ipa_ale_df$bClass=="IPA"])
t.test(ipa_ale_df$ABV[ipa_ale_df$bClass=="Ale"])
```
